Evaluating the Effects of Tracker Reliability and Field of View on a Target Following Task in Augmented Reality

### 2009-06-25 Paper/paper.tex
added text about display options

We developed a target following task scenario, where the user is asked to visually track a virtual person as it moves throughout the scene.  The participant is centered inside a room with various doors and windows giving a view to the outside world.  They may change their orientation as the please, but not change position (3 degrees of freedom).  The virtual person to be tracked walks an unpredictable path outside the room.  The user's view to the virtual person may be occluded at any time by the walls of the building; at other times the virtual person may be visible through doors or windows in the building.  The augmented view element is composed of red marker rectangles indicating the location of the tracked person and other virtual people.

Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.  The augmented overlays are merged with the original image to form the simulated augmented reality view.

The parameters of our experiment are divided between the ``real'' components of immersion, such as the field of view of the HMD, and the ``augmented'' components of immersion, such as the field of view of the simulated AR display, and the performance of the tracking sensor.  We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.  For example, such dropouts might occur with a magnetic sensor near interfering material.

@@

We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.  Figure \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view have a very significant effect on performance using this metric.  
% TODO: interaction?

We also used a two-way {ANOVA} to examine the number of times the participant loses the target, with results shown in Figure \ref{fig:anova}.  Both field of view and dropout length have a significant effect.  There also is an interaction between the two factors.  Figure \ref{fig:interaction} shows a plot of this interaction.

>>>>

We developed a target following task scenario, where the user is asked to visually track a virtual person as it moves throughout the scene.  The participant is centered inside a room with various doors and windows giving a view to the outside world.  They may change their orientation, but not change position (3 degrees of freedom).  The user's view to the virtual people may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.  The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and distractor people.

The parameters of our experiment are divided between the ``real'' components of immersion, such as the field of view of the HMD, and the ``augmented'' components of immersion, such as the field of view of the simulated AR display, and the performance of the tracking sensor.  We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.  For example, such dropouts might occur with a magnetic sensor near interfering material.  Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.  However, these options could have a positive or negative effect on performance, depending on the circumstances.  We determined that hiding the AR overlay during sensor dropouts has the least effect on performance.

@@

We performed two-way {ANOVA} for the time following target metric as an overall analysis of the effects of our two factors.  Figure \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view have a very significant effect on performance using this metric.  
% TODO: interaction?

We also used a two-way {ANOVA} to examine the number of times the participant loses the target, with results shown in Figure \ref{fig:anova}.  Both field of view and dropout length have a significant effect.  There also is an interaction between the two factors.  Figure \ref{fig:interaction} shows a plot of this interaction.  
% TODO: need F values reported here.
